A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2018-01-30 paper.tex
update implementation section, use minted for code snippets

%%%python syntax highlighting
% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{10} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{10}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

\usepackage{listings}

% Python style for highlighting
\newcommand\pythonstyle{\lstset{
language=Python,
basicstyle=\ttm,
otherkeywords={self},             % Add keywords here
keywordstyle=\ttb\color{deepblue},
emph={MyClass,__init__},          % Custom highlighting
emphstyle=\ttb\color{deepred},    % Custom highlighting style
stringstyle=\color{deepgreen},
frame=tb,                         % Any extra options here
showstringspaces=false            % 
}}

% Python environment
\lstnewenvironment{python}[1][]
{
\pythonstyle
\lstset{#1}
}
{}

@@

The \texttt{HarmonizationDriver} is an object designed to interface with
user-provided data and configuration files. Input data is assumed to be an Excel
file the in the in the IAMC format, i.e., using \code{Model}, \code{Scenario},
\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
representing each modeled time period, which is the standard format within the
IAM community. It is responsible for downselecting data into unique combinations
for each model and scenario and then invoking the harmonization
process. Finally, it adds any exogenous emissions trajectories as required by
the user and recompiles the results and writes them to disc. The
\texttt{HarmonizationDriver} acts the primary interface for high-level users as
shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided

\begin{python}[caption=High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}, label=lst:tutorial]
from aneris.tutorial import load_data
model, hist, driver = load_data()
for scenario in driver.scenarios():
    driver.harmonize(scenario)
harmonized, metadata = driver.harmonized_results()
\end{python}

\begin{lstlisting}[language=bash, caption=bash, label=lst:cli]
$ aneris -h
usage: aneris [-h] [--history HISTORY] [--regions REGIONS] [--rc RC]
              [--output_path OUTPUT_PATH] [--output_prefix OUTPUT_PREFIX]
              input_file
    Harmonize historical trajectories to data in the IAMC template format.
    Example usage:
    aneris input.xlsx --history history.csv --regions regions.csv
positional arguments:
  input_file            Input data file.
optional arguments:
  -h, --help            show this help message and exit
  --history HISTORY     Historical emissions in the base year.
  --regions REGIONS     Mapping of country iso-codes to native regions.
  --rc RC               Runcontrol YAML file (see
                        http://mattgidden.com/aneris/config.html for
                        examples).
  --output_path OUTPUT_PATH
                        Path to use for output file names.
  --output_prefix OUTPUT_PREFIX
                        Prefix to use for output file names.
\end{lstlisting}

@@

\begin{figure}
  \subfloat[Python Implementation\label{fig:software}]{
    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
    }
  \subfloat[Workflow\label{fig:workflow}]{
    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
    }
    \caption{
      The full harmonization process as executed by \code{aneris}. Operations
      that can be configured with user-based input configurations are shown in
      purple. The core harmonization process is shown in yellow.  }
    \label{fig:software-workflow}
\end{figure}

Users are able to control the harmonization process via a number of options
(with examples provided
\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
by which users control the process is by providing \textit{override} methods for
any combination of region and variable (i.e., sector and gas species).  In
practice, it may be possible that not all default methods chosen will provide
robust harmonized trajectories, especially if there is a significant difference
between historical and model values in the harmonization year, if there is
significant upward or downward movement in the model trajectory, or if there are
known discrepancies in sectoral definition between the IAM and historical data
source. In such cases, override methods can ameliorate the issues associated
with the default method choice. In order to help identify these cases,
harmonization \textit{diagnostics} are provided which analyze the relative
difference between harmonized and unharmonized trajectories at their mid (if $>
400$\%) and end-points (if $> 200$\%). If override methods are provided, they
are used instead of the default methods as determined by the default method
decision tree. Furthermore, users can set the above-mentioned thresholds as well
as the LUC method used in the decision tree (see Figure
\ref{fig:decision_tree}).

>>>>

@@

The \texttt{HarmonizationDriver} is an object designed to interface with
user-provided data and configuration files. Input data is assumed to be an Excel
file the in the in the IAMC format, i.e., using \code{Model}, \code{Scenario},
\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
representing each modeled time period, which is the standard format within the
IAM community. It is responsible for downselecting data into unique combinations
for each model and scenario and then invoking the harmonization
process. Finally, it adds any exogenous emissions trajectories as required by
the user and recompiles the results and writes them to disc. The
\texttt{HarmonizationDriver} acts the primary interface for high-level users as
shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to
allow users to more easily incorporate the harmonization process in scripted
workflows (Listing \ref{lst:cli}, Figure \ref{fig:software}).

\begin{listing}[H]
\begin{minted}{python}
from aneris.tutorial import load_data
model, hist, driver = load_data()
for scenario in driver.scenarios():
    driver.harmonize(scenario)
harmonized, metadata = driver.harmonized_results()
\end{minted}
\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}}
\label{lst:tutorial}
\end{listing}

\begin{listing}[H]
\begin{minted}{bash}
$ aneris -h
usage: aneris [-h] [--history HISTORY] [--regions REGIONS] [--rc RC]
              [--output_path OUTPUT_PATH] [--output_prefix OUTPUT_PREFIX]
              input_file

    Harmonize historical trajectories to data in the IAMC template format.

    Example usage:

    aneris input.xlsx --history history.csv --regions regions.csv

positional arguments:
  input_file            Input data file.

optional arguments:
  -h, --help            show this help message and exit
  --history HISTORY     Historical emissions in the base year.
  --regions REGIONS     Mapping of country iso-codes to native regions.
  --rc RC               Runcontrol YAML file (see
                        http://mattgidden.com/aneris/config.html for
                        examples).
  --output_path OUTPUT_PATH
                        Path to use for output file names.
  --output_prefix OUTPUT_PREFIX
                        Prefix to use for output file names.
\end{minted}
\caption{The \gls{cli} help provided by the \code{aneris} package.}
\label{lst:cli}
\end{listing}

@@

\begin{figure}
  \subfloat[Python Implementation\label{fig:software}]{
    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
    }
  \subfloat[Workflow\label{fig:workflow}]{
    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
    }
    \caption{ Panel \ref{fig:software} shows the various objects in the
      \code{aneris} codebase as well as a short description of their scope of
      concern. Panel \ref{fig:workflow} displays the full harmonization process
      as executed by \code{aneris} for a single instance of a model and
      scenario. Operations that can be configured with user-based input
      configurations are shown in purple. Operations goverened by the
      \texttt{HarmonizationDriver} are shown in red. Data processing operations
      are shown in blue. The core harmonization process, governed by the
      \texttt{Harmonizer} is shown in green.  }
    \label{fig:software-workflow}
\end{figure}

Users are able to control the harmonization process via a number of options
(with examples provided
\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
by which users control the process is by providing \textit{override} methods for
any combination of region and variable (i.e., sector and gas species).  In
practice, it may be possible that not all default methods chosen will provide
robust harmonized trajectories, especially if there is a significant difference
between historical and model values in the harmonization year, if there is
significant upward or downward movement in the model trajectory, or if there are
known discrepancies in sectoral definition between the IAM and historical data
source. In such cases, override methods can ameliorate the issues associated
with the default method choice.

In order to help identify these cases, harmonization \textit{diagnostics} are
provided which analyze the relative difference between harmonized and
unharmonized trajectories at their mid and end-points. These values are
configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
provided based on experiences of using \code{aneris} to date. If override
methods are provided, they are used instead of the default methods as determined
by the default method decision tree. Furthermore, users can set the
above-mentioned thresholds as well as the LUC method used in the decision tree
(see Figure \ref{fig:decision_tree}).
